Intrinsic Verification of Parsers and Formal Grammar Theory in Dependent Lambek Calculus

### 2024-10-24 paper/paper.tex
fixed equalizer command
equalizer intro, beta, eta
Some examples

% -*- fill-column: 80; -*-
\documentclass[sigconf,review,anonymous,screen,acmsmall]{acmart}
\usepackage{mathpartir}
\usepackage{tikz-cd}
\usepackage{lipsum}
\usepackage{enumitem}
\usepackage{wrapfig}
\usepackage{fancyvrb}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{scalerel}
\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows, fit}

@@

\newcommand{\SPF}{\mathsf{SPF}}
\newcommand{\Var}{\mathsf{Var}}
\newcommand{\K}{\mathsf{K}}
\newcommand{\map}{\mathsf{map}}
\newcommand{\roll}{\mathsf{roll}}
\newcommand{\fold}{\mathsf{fold}}
\newcommand{\sem}[1]{\llbracket{#1}\rrbracket}
\newcommand{\cat}[1]{\mathbf{#1}}
\newcommand{\lto}{\multimap}
\newcommand{\tol}{\mathrel{\rotatebox[origin=c]{180}{$\lto$}}}
\newcommand{\String}{\Sigma^{*}}
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Syn}{\mathbf{Synx}}
\newcommand{\SemAct}{\mathbf{SemAct}}
\newcommand{\Gr}{\mathbf{Gr}}
\newcommand{\Grammar}{\mathbf{Gr}}
\newcommand{\Type}{\mathbf{Type}}
\newcommand{\Prop}{\mathbf{Prop}}
\newcommand{\Bool}{\mathbf{Bool}}
\newcommand{\true}{\mathbf{true}}
\newcommand{\false}{\mathbf{false}}
\newcommand{\nat}{\mathbb{N}}
\newcommand{\theoryname}{Dependent Lambek Calculus\xspace}
\newcommand{\theoryabbv}{$\textrm{Lambek}^D$}

@@

@@

@@

\section{Dependent Lambek Calculus by Example}
\label{sec:type-theory-examples}

@@

@@

    \inferrule{~}{\Gamma ; a : A \vdash a : A}
    \and
    \inferrule{\Gamma ; \Delta \vdash e : B \\ \linctxwffjdg \Gamma {A \equiv B}}{\Gamma ; \Delta \vdash e : A}
    %
    \\
    %
    \inferrule{~}{\Gamma ; \cdot \vdash () : I}
    \and
    \inferrule{\Gamma ; \Delta \vdash e : I \\ \Gamma ; \Delta_1',\Delta_2' \vdash e' : C}{\Gamma ; \Delta_1',\Delta,\Delta_2' \vdash \letin {()} e {e'} : C}
    %
    \\
    %
    \inferrule{\Gamma ; \Delta \vdash e : A \\ \Gamma ; \Delta' \vdash e' : B}{\Gamma ; \Delta, \Delta' \vdash e \otimes e' : A \otimes B}
    %
    \and
    %
    \inferrule{\Gamma ; \Delta \vdash e : A \otimes B \\ \Gamma ; \Delta'_1, a : A, b : B, \Delta'_2 \vdash e'}{\Gamma ;  \Delta_1', \Delta, \Delta'_2 \vdash \letin {a \otimes b} e {e'}}
    \\
    %
    \inferrule{\Gamma ; a : A , \Delta \vdash e : B}{\Gamma ; \Delta \vdash \lamblto a e : A\lto B}
    \and
    \inferrule{\Gamma ; \Delta' \vdash e' : A \\ \Gamma ; \Delta \vdash e : A \lto B}{\Gamma ; \Delta', \Delta \vdash \applto {e'} {e} : B}
    \\
    %
    \inferrule{\Gamma ; \Delta , a : A \vdash e : B}{\Gamma ; \Delta \vdash \lambtol a e : B\tol A}
    \and
    \inferrule{\Gamma ; \Delta \vdash e : B \tol A \\ \Gamma ; \Delta' \vdash e' : A}{\Gamma ; \Delta, \Delta' \vdash \apptol e {e'} : B}
    %
    \\
    %
    \inferrule{\Gamma, x : X ; \Delta  \vdash e : A}
              {\Gamma ; \Delta \vdash \dlamb x e : \LinPiTy x X A}
    \and
    \inferrule{\Gamma ; \Delta \vdash e : \LinPiTy x X A \\ \Gamma \vdash e' : X}{\Gamma ; \Delta \vdash e\,.\pi\,M : \subst A {e'} x}
    %
    \\
    %
    \inferrule{\Gamma \vdash M : X \quad \Gamma ; \Delta \vdash e : \subst A e x}{\Gamma ; \Delta \vdash \sigma\,M\,e : \bigoplus\limits_{x:X} A}
    %
    \and
    %
    \inferrule{\Gamma ; \Delta \vdash e : \bigoplus\limits_{x:X} A \quad \Gamma, x : X ; \Delta'_1, a : A, \Delta'_2 \vdash e' : C}{\Gamma; \Delta'_1, \Delta, \Delta'_2 \vdash \letin {\sigma\,x\,a} e {e'}: C}
    %% %
    %% \\
    %% %
    %% \inferrule{~}{\Gamma ; \Delta \vdash () : \top}

@@

\begin{figure}
  \begin{mathpar}
    \begin{array}{rcl}
      \textrm{non-linear types} X,Y,Z & ::= & \textrm{Empty}\pipe\textrm{Unit}\pipe\textrm{Bool}\pipe\textrm{Nat}
      \pipe \sum_{x:X} Y \pipe \prod_{x:X}Y\pipe \textrm{SPF}\,X \pipe \cdots \max{also eliminators...}\\
      \textrm{linear types} A,B,C & ::= & c \pipe I \pipe A \otimes B \pipe A \lto B \pipe B \tol A
      \pipe \bigoplus_{x:X} A \pipe \bigamp_{x:X} A \pipe \mu e e'\\
      \textrm{non-linear terms} M,N,O & ::= & \cdots\\
      \textrm{linear terms} 
    \end{array}
  \end{mathpar}
  \caption{Syntax of Types and Terms}
\end{figure}

@@

    \and
    \inferrule{c \in \Sigma}{\Gamma \vdash c : L_0}
    %
    \and
    %
    \inferrule{\Gamma, x : L_i \vdash A : L_i \and A \textrm{ strictly positive}}{\Gamma \vdash \mu x.\, A : L_i}
  \end{mathpar}
  \caption{Type well-formedness}
  \label{fig:typewf}
\end{figure}

@@

    \inferrule{~}{\Gamma; \cdot \vdash G^{-1}\, (G \, t ) \equiv t: A}
%
    \\
%
    \inferrule{~}{\Gamma; \Delta \vdash \app {(\lamblto a e)} {e'} \equiv \subst e x {e'} : C}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash e \equiv \lamblto a {\app e a} : A \lto B}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash \app {(\lambtol a e)} {e'} \equiv \subst e x {e'} : C}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash e \equiv \lambtol a {\app e a} : A \tol B}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash \app {(\dlamb x e)} {e'} \equiv \subst x a {e'} : C}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash e \equiv \dlamb x {\app e x} : \LinPiTy x X A}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash e \equiv e' : \top}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash e_i \equiv \pi_i (e_1, e_2) : A_i}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash e \equiv (\pi_1 e, \pi_2 e) : A\& B}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash \letin {()} {()} e \equiv e : C}
    %
    \and
    %
    \inferrule{~}{\Gamma; \Delta \vdash \letin {()} e {\subst {e'} {()} a} \equiv \subst {e'} a e : C}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash \letin {e \otimes e'} {a \otimes a'} e'' \equiv \subst {e''} {a, a'} {e, e'} : C}
%
    \and
    %
    \inferrule{~}{\Gamma; \Delta \vdash \letin {a \otimes b} e {\subst {e'} {a \otimes b} c} \equiv \subst {e'} c e : C}
%
    \and
%
    \inferrule{~}{\Gamma;\Delta \vdash \letin {(x, a)} {(e, e')} {e''} \equiv \subst {e''} {x, a} {e, e'} : C}
    %
    \and
    %
    \inferrule{~}{\Gamma; \Delta \vdash \letin {(x, a)} e {\subst {e'} {(x, a)} y} \equiv \subst {e'} e y : C}
\end{mathpar}
  \caption{Judgmental equality}
  \label{fig:jdgeq}
\end{figure}

>>>>

% -*- fill-column: 80; -*-
\documentclass[sigconf,review,anonymous,screen,acmsmall]{acmart}
\usepackage{mathpartir}
\usepackage{tikz-cd}
\usepackage{lipsum}
\usepackage{enumitem}
\usepackage{wrapfig}
\usepackage{fancyvrb}
\usepackage{xspace}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{scalerel}
\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows, fit}

@@

\newcommand{\SPF}{\mathsf{SPF}}
\newcommand{\Var}{\mathsf{Var}}
\newcommand{\K}{\mathsf{K}}
\newcommand{\map}{\mathsf{map}}
\newcommand{\roll}{\mathsf{roll}}
\newcommand{\fold}{\mathsf{fold}}
\newcommand{\inl}{\mathsf{inl}}
\newcommand{\inr}{\mathsf{inr}}
\newcommand{\sem}[1]{\llbracket{#1}\rrbracket}
\newcommand{\cat}[1]{\mathbf{#1}}
\newcommand{\lto}{\multimap}
\newcommand{\tol}{\mathrel{\rotatebox[origin=c]{180}{$\lto$}}}
\newcommand{\String}{\Sigma^{*}}
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Syn}{\mathbf{Synx}}
\newcommand{\SemAct}{\mathbf{SemAct}}
\newcommand{\Gr}{\mathbf{Gr}}
\newcommand{\Grammar}{\mathbf{Gr}}
\newcommand{\Type}{\mathbf{Type}}
\newcommand{\Prop}{\mathbf{Prop}}
\newcommand{\Bool}{\mathbf{Bool}}
\newcommand{\true}{\mathbf{true}}
\newcommand{\false}{\mathbf{false}}
\newcommand{\nat}{\mathbb{N}}
\newcommand{\theoryname}{Dependent Lambek Calculus\xspace}
\newcommand{\theoryabbv}{$\textrm{Lambek}^D$}

@@

\newcommand{\bang}{~\textbf{!}~}
\newcommand{\nil}{\mathsf{nil}}
\newcommand{\cons}{\mathsf{cons}}
\newcommand{\data}{\mathsf{data}}
\newcommand{\stringquote}[1]{\texttt{\textquotesingle#1\textquotesingle}}

@@

\newcommand{\equalizer}[3]{\{#1\,|\,\apptol {#2}{#1} = \apptol{#3}{#1} \}}
\newcommand{\equalizerpair}[2]{\langle #1 , #2 \rangle}
\newcommand{\equalizerpi}[1]{#1.\pi}

@@

\section{\theoryname by Example}
\label{sec:type-theory-examples}

To gain intuiton for working in \theoryname, let's walk through some examples of
grammars and parses internal to it.

\paragraph{Finite Grammars}
First, consider finite grammars over the alphabet $\{ a , b , c \}$. Each letter of the
alphabet corresponds to a base type in our calculus. For instance, the grammar
for the single character $a$ has a single parse tree for the string \stringquote{a}.
and no parse trees at any other strings.

A finite grammar is built from base types via dijunctions $\oplus$ and
concatenations $\otimes$ --- for instance, $(a \oplus b) \otimes c$.
In \cref{fig:fingram} we find a proof tree showing that the input string \stringquote{ac} matches
the grammar $(a \oplus b) \otimes c$.

\begin{figure}
\begin{mathpar}
  \inferrule
  {
    \inferrule
    {~}
    {x : a \vdash \inl(x) : a \oplus b}
    \\
    \inferrule
    {~}
    {y : c \vdash y : c}
  }
  {x : a , y : c \vdash \inl(x) \otimes y : (a \oplus b) \otimes c}
\end{mathpar}
\caption{\stringquote{ac} matches $(a \oplus b) \otimes c$}
\label{fig:fingram}
\end{figure}

To parse the specific string \stringquote{ac}, we populate the context with an
internalized encoding of that string. In contexts of this form, linear terms
describe the structure of a parse tree. That is, $\inl(x) \otimes y$ details the
manner by which \stringquote{ac} matches $(a \oplus b) \otimes c$.

Because derivations in a linear context correspond to parse trees in this
manner, we cannot affirm all of the usual structural rules of a type system.
Weakening and contraction each allow for modification of the string in context ---
weakening allows for characters to be dropped at will, while contraction allows
for duplication of characters. Omission of these rules offers a
resource-sensitive view that ensures only the inputted characters are parsed.
Moreover, strings are not simply sets of characters, as they carry an inherent
ordering. Thus, the rule of exchange is also omitted in this calculus so that
the constituent characters of a string are not freely permuted.

Without the above structural rules we
arrive at an ordered, linear type system, and to summarize this
particular choice of substructural restrictions: when parsing, characters must
be processed
without modification and in the order in which they were written.

\paragraph{Regular Expressions}
Next, we demonstrate how to parse a regular expression in \theoryabbv. Again,
consider grammars over $\{ a , b , c \}$. Similar to finite grammars, regular
expresssions are built inductively using base types, $\oplus$, and $\otimes$ ;
however, unlike finite grammars, regular expressions additionally have a
restricted form of recursion via Kleene star $(\cdot)^{\ast}$.

% TODO this should be beter formatted but can't use listings with \otimes ^{\ast}
\begin{figure}
\begin{align*}
\data &~A^{*}:\\
      & \nil : I \to A* \\
      & \cons : A \otimes A* \to A*
\end{align*}
\caption{Kleene Star as an inductive type}
\label{fig:kleenestarinductive}
\end{figure}

For a grammar $A$, the recursive nature of $A^{*}$ is captured by defining
$A^{*}$ to be an inductive linear
type, as shown in \cref{fig:kleenestarinductive}. Just as in \cref{fig:fingram}, we can show that particular strings match a Kleene
star. In \cref{fig:kleenestarderivation}, we demonstrate that \stringquote{ab}
matches $a^{*} \otimes b$ through repeated application of the Kleene star constructors.

\begin{figure}
\begin{mathpar}
  \inferrule
  {
    \inferrule
    {
      \inferrule
      {~}
      {x : a \vdash x : a}
      \\
      \inferrule
      {
        \inferrule
        {~}
        {I \vdash \nil : a^{*}}
      }
      {\cdot \vdash \nil : a^{*}}
    }
    {x : a \vdash \cons (x , \nil) : a^{*}}
    \\
    \inferrule
    {~}
    {y : b \vdash y : b}
  }
  {x : a , y : b \vdash \cons(x ) \otimes y : a^{*} \otimes b}
\end{mathpar}
\caption{\stringquote{ab} matches $a^{*} \otimes b$}
\label{fig:kleenestarderivation}
\end{figure}

However, we can have derivations where the term in context is not simply a
string of base types. For instance, by using the elimination principle for
Kleene star, \cref{fig:kleeneabstractproof} builds a term of type
$A^{*}$ in context comprising abstract\footnote{A linear type that is not a
  finite grammar.} linear types. In these contexts, it is not quite appropriate
to treat linear terms as parse \textit{trees}, because the information in
context is not a string. Instead, it is more
appropriate to think of such terms as parse \textit{transformers}. That is, in
\cref{fig:kleeneabstractproof} the
inhabitant $x : (A \otimes A)^{*}$ in the context is not itself data to be parsed; rather,
$x$ is a parse tree and the term from context $x : (A \otimes A)^{*}$ to the
term $A^{*}$ is a transformation of parse trees of the input grammar into parse
trees of the output grammar.

\steven{TODO : describe elimination principle somewhere. Maybe just in prose?}
\steven{I probably can't get away with the inline definition of $f$ below}
\begin{figure}
\begin{mathpar}
  \inferrule
  {
    \inferrule
    {~}
    {x_I : I \vdash \nil : A^*}
    \\
    \inferrule
    {
      \inferrule
      {~}
      {x_1 : A , x_2 : A , y : A* \vdash \cons (x_1 , \cons(x_2 , y)) : A^*}
    }
    {x : A \otimes A , y : A^*  \vdash f := \letin {x_1 \otimes x_2} {x} {\cons(x_1 ,
        \cons(x_2 , y))} : A^*}
  }
  {x : (A \otimes A)^* \vdash \fold(\nil , f) : A^* }
\end{mathpar}
\caption{An example mapping out of abstract grammars}
\label{fig:kleeneabstractproof}
\end{figure}

% \begin{tikzpicture}[%opacity=0.5,
%   treenode/.style = {shape=rectangle, rounded corners,
%     draw, align=center,
%     top color=white, bottom color=white!20},
%     line/.style={draw, -latex'},
%     edge from parent/.style={draw,-latex'},
%     level 1/.style={sibling distance=20mm, level distance=0.5cm},
%     level 2/.style={sibling distance=1cm, level distance=1cm},
%     visible on=<3->
%     ]
%   \node[treenode] (tree) {$\otimes$}
%     child { node[treenode] (a) {${\color{blue} a}$} }
%     child { node[treenode] (b){${\color{magenta} \mathsf{inl}}$}
%         child { node[treenode] (c) {${\color{magenta} b}$} }
%       };
%       \node[draw, fit=(tree)(a)(b)(c),
%       label={[name=l] $g(ab)$}, visible on=<3->] {};
%     \node[below=2.5cm of tree] (mt) {};
%     \node[below=2.5cm of a] (mta) {};
%     \node[below=2.5cm of b] (mtb) {};
%     \node[below=-0.2cm of mt, visible on=<4->] (mtset) {};
%       \node[draw, fit=(mt)(mtset)(mta)(mtb),
%       label={[name=l, visible on=<4->] $g(aa)$}, visible on=<4->] {};
% \end{tikzpicture}

@@

\steven{This figure is broken}
\begin{figure}
  % \begin{array}{rrcl}
  %   \textrm{non-linear contexts} & \Gamma & ::= & \cdot \pipe \Gamma,x:X \\
  %   \textrm{non-linear types} & X , Y , Z & ::= & \mathsf{Nat} \pipe
  %                                                 \mathsf{Int} \pipe
  %                                                 \mathsf{Empty} \pipe
  %                                                 \mathsf{Unit} \pipe
  %                                                 \mathsf{SPF}\,X\pipe U \pipe L
  %                                                 \pipe !A \pipe \sum_{x:X} Y
  %                                                 \pipe \prod_{x:X} Y\pipe M =_A
  %                                                 N \\
  %   \textrm{linear contexts} & \Delta & ::= & \cdot \pipe \Delta , a:A\\
  %   \textrm{linear types} & A , B , C & ::= &
  %   c \pipe I \pipe A \otimes B \pipe \bigoplus\limits_{x:X} A \pipe \mu A_f M
  %   A \lto B \pipe B \tol A \pipe \bigwith\limits_{x:X} A \pipe \equalizer {a}{f}{g}\\
  % \end{array}
  \caption{Overview of Types}
\end{figure}

@@

    \inferrule{~}{\Gamma ; a : A \vdash a : A}
    \and
    \inferrule{\Gamma ; \Delta \vdash e : B \\ \linctxwffjdg \Gamma {A \equiv B}}{\Gamma ; \Delta \vdash e : A}
    %
    \\
    %
    \inferrule{~}{\Gamma ; \cdot \vdash () : I}
    \and
    \inferrule{\Gamma ; \Delta \vdash e : I \\ \Gamma ; \Delta_1',\Delta_2' \vdash e' : C}{\Gamma ; \Delta_1',\Delta,\Delta_2' \vdash \letin {()} e {e'} : C}
    %
    \\
    %
    \inferrule{\Gamma ; \Delta \vdash e : A \\ \Gamma ; \Delta' \vdash e' : B}{\Gamma ; \Delta, \Delta' \vdash e \otimes e' : A \otimes B}
    %
    \and
    %
    \inferrule{\Gamma ; \Delta \vdash e : A \otimes B \\ \Gamma ; \Delta'_1, a : A, b : B, \Delta'_2 \vdash e'}{\Gamma ;  \Delta_1', \Delta, \Delta'_2 \vdash \letin {a \otimes b} e {e'}}
    \\
    %
    \inferrule{\Gamma ; a : A , \Delta \vdash e : B}{\Gamma ; \Delta \vdash \lamblto a e : A\lto B}
    \and
    \inferrule{\Gamma ; \Delta' \vdash e' : A \\ \Gamma ; \Delta \vdash e : A \lto B}{\Gamma ; \Delta', \Delta \vdash \applto {e'} {e} : B}
    \\
    %
    \inferrule{\Gamma ; \Delta , a : A \vdash e : B}{\Gamma ; \Delta \vdash \lambtol a e : B\tol A}
    \and
    \inferrule{\Gamma ; \Delta \vdash e : B \tol A \\ \Gamma ; \Delta' \vdash e' : A}{\Gamma ; \Delta, \Delta' \vdash \apptol e {e'} : B}
    %
    \\
    %
    \inferrule{\Gamma, x : X ; \Delta  \vdash e : A}
              {\Gamma ; \Delta \vdash \dlamb x e : \LinPiTy x X A}
    \and
    \inferrule{\Gamma ; \Delta \vdash e : \LinPiTy x X A \\ \Gamma \vdash e' : X}{\Gamma ; \Delta \vdash e\,.\pi\,M : \subst A {e'} x}
    %
    \\
    %
    \inferrule{\Gamma \vdash M : X \quad \Gamma ; \Delta \vdash e : \subst A e x}{\Gamma ; \Delta \vdash \sigma\,M\,e : \bigoplus\limits_{x:X} A}
    %
    \and
    %
    \inferrule{\Gamma ; \Delta \vdash e : \bigoplus\limits_{x:X} A \quad \Gamma, x : X ; \Delta'_1, a : A, \Delta'_2 \vdash e' : C}{\Gamma; \Delta'_1, \Delta, \Delta'_2 \vdash \letin {\sigma\,x\,a} e {e'}: C}
    %
    \and
    %
    \inferrule
    {~}
    {\Gamma , x : \square A ; \cdot \vdash \bang x : A}
    %
    \and
    %
    \inferrule
    {\Gamma \vdash f : \square (A \lto B) \\
      \Gamma \vdash g : \square (A \lto B) \\
      \Gamma ; \Delta \vdash a : A \\
      \Gamma ; \Delta \vdash p : \apptol {\bang f}{a} = \apptol {\bang g}{a}}
    {\Gamma ; \Delta \vdash \equalizerpair{a}{p} : \equalizer {a}{\bang f}{\bang
      g}}
    %
    \and
    %
    \inferrule
    {\Gamma \vdash f : \square (A \lto B) \\
      \Gamma \vdash g : \square (A \lto B) \\
      \Gamma ; \Delta \vdash e : \equalizer{a}{\bang f}{\bang g}}
    {\Gamma ; \Delta \vdash \equalizerpi {e} : A}
    %% \\
    %% %
    %% \inferrule{~}{\Gamma ; \Delta \vdash () : \top}

@@

@@

    \and
    \inferrule{c \in \Sigma}{\Gamma \vdash c : L_0}
    %
    \and
    %
    \inferrule{\Gamma, x : L_i \vdash A : L_i \and A \textrm{ strictly positive}}{\Gamma \vdash \mu x.\, A : L_i}
    %
    \and
    %
    \inferrule
    {\Gamma \vdash A : L_i}
    {\Gamma \vdash \square A : U_i}
    %
    \and
    \inferrule
    {\Gamma ; \Delta \vdash f : A \\ \Gamma \vdash g : \square (A \lto B)}
    {\Gamma ; \Delta \vdash \equalizer {a}{\bang f}{\bang g} : L_i}
  \end{mathpar}
  \caption{Type well-formedness}
  \label{fig:typewf}
\end{figure}

@@

    \inferrule{~}{\Gamma; \cdot \vdash G^{-1}\, (G \, t ) \equiv t: A}
%
    \\
%
    \inferrule{~}{\Gamma; \Delta \vdash \app {(\lamblto a e)} {e'} \equiv \subst e x {e'} : C}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash e \equiv \lamblto a {\app e a} : A \lto B}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash \app {(\lambtol a e)} {e'} \equiv \subst e x {e'} : C}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash e \equiv \lambtol a {\app e a} : A \tol B}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash \app {(\dlamb x e)} {e'} \equiv \subst x a {e'} : C}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash e \equiv \dlamb x {\app e x} : \LinPiTy x X A}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash e \equiv e' : \top}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash e_i \equiv \pi_i (e_1, e_2) : A_i}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash e \equiv (\pi_1 e, \pi_2 e) : A\& B}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash \letin {()} {()} e \equiv e : C}
    %
    \and
    %
    \inferrule{~}{\Gamma; \Delta \vdash \letin {()} e {\subst {e'} {()} a} \equiv \subst {e'} a e : C}
%
    \and
%
    \inferrule{~}{\Gamma; \Delta \vdash \letin {e \otimes e'} {a \otimes a'} e'' \equiv \subst {e''} {a, a'} {e, e'} : C}
%
    \and
    %
    \inferrule{~}{\Gamma; \Delta \vdash \letin {a \otimes b} e {\subst {e'} {a \otimes b} c} \equiv \subst {e'} c e : C}
%
    \and
%
    \inferrule{~}{\Gamma;\Delta \vdash \letin {(x, a)} {(e, e')} {e''} \equiv \subst {e''} {x, a} {e, e'} : C}
    %
    \and
    %
    \inferrule{~}{\Gamma; \Delta \vdash \letin {(x, a)} e {\subst {e'} {(x, a)} y} \equiv \subst {e'} e y : C}
    \and
    %
    \inferrule
    {\Gamma \vdash f : \square (A \lto B) \\
      \Gamma \vdash g : \square (A \lto B) \\
      \Gamma ; \Delta \vdash a : A \\
      \Gamma ; \Delta \vdash p : \apptol {\bang f}{a} \equiv \apptol {\bang g}{a}}
    {\Gamma ; \Delta \vdash \equalizerpi {\equalizerpair {a}{p}} \equiv a : A}
    %
    \and
    %
    \inferrule
    {\Gamma \vdash f : \square (A \lto B) \\
      \Gamma \vdash g : \square (A \lto B) \\
      \Gamma ; \Delta \vdash p : \apptol {\bang f}{a} = \apptol {\bang g}{a} \\
      \Gamma ; \Delta \vdash e : \equalizer{a}{\bang f}{\bang g}}
    {\Gamma ; \Delta \vdash \equalizerpair {\equalizerpi {e}} {p} \equiv e : A}
\end{mathpar}
  \caption{Judgmental equality}
  \label{fig:jdgeq}
\end{figure}
